\documentclass{article}
\usepackage{amsmath,amsfonts,amsthm,amssymb}
\usepackage{siunitx}
\usepackage{graphicx,float,wrapfig}
\usepackage{cite}

\sisetup{per-mode=symbol}

\author{Guilherme Nettesheim}
\title{Experiment results}

\graphicspath{ {../images/} }

\begin{document}
\maketitle

\section{February 14-17, 2019}

Summary
\begin{description}
\item[Date:] 12:30pm Thursday, February 14 2019 to 11:50pm Sunday, February 17 2019.
\item[Duration:] 83 hours and 20 minutes, 12000 frames
\item[Purpose:] Observe the sequence egress $\to$ invasion $\to$ egress to quantify malaria growth times and expression of the fluorescent, apical protein
\item[Data location:] External hard-drive malaria1, \texttt{malaria1/data/14-17\_february\_experiment}. Includes segmented data, \texttt{.temperature}, and \texttt{.multiled} files
\item[Summary:] Most of trophozoites egressed successfully. Only one egress led to an invasion. Bad autofluorescence by end of experiment.  
\end{description}


Sample preparation
\begin{description}
\item[Malaria:] RON11 3D7 GFP
\item[Parasitaemia:] 2\% early trophozoite, 1\% rings at 8:30am, four hours before start of experiment
\item[Sample haematocrit:] \SI{20}{\micro\liter} of 50\% Hct in \SI{5}{\milli\liter} growth medium. Used \SI{4}{\milli\liter}.
\item[Surface sticking:] Incubated \SI{1}{\milli\liter} of \SI{5}{\milli\gram\per\milli\liter} Concanavalin A for over 30 minutes. Washed out three times with wash medium.
\item[Microscope:] Right microscope in POM 0.26
\item[Heating:] .
	\begin{description}
	\item[Chamber:] Set to \SI{37}{\celsius} using feedback
	\item[Lid:] Set to \SI{38}{\celsius} using a constant voltage suppy at \SI{5.5}{\volt}, \SI{160}{\milli\ampere}
	\item[Collar:] Set to \SI{37}{\celsius} using feedback, however temperature was measured at collar with NTC connection, and not a thermocouple near objective
	\end{description}
\item[Gas:] Mixture of 1\% O$_2$, 3\%CO$_2$, N$_2$ balance flowed at \SI{1}{\liter\per\minute} at \SI{2.8}{\bar}.
\item[Flow:] Growth medium was pumped through at \SI{10}{\micro\liter\per\minute}
\end{description}

Imaging settings
\begin{description}
	\item[Lens:] Plan Apo 60X/1.40 oil $\infty$/0.17 DIC N2
	\item[Exposure time:] Roughly \SI{20}{\milli\second}
	\item[Fluorescence dosage:] \SI{20}{\milli\second} of blue light every \SI{5}{\minute} at 0.9 intensity, or \SI{216}{\milli\second\per\hour} intensity.
	\item[Brightfield dosage:] \SI{20}{\milli\second} of red light every \SI{5}{\minute} at 0.01 intensity, or \SI{2.4}{\milli\second\per\hour} intensity.
\end{description}

\end{document}
